---
title: 'Intro to Bayesian Statistic: Simple Regression Analysis' 
author: "Philipp Masur"
date: "2022-09"
output: 
  github_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r, echo=F, message=F}
knitr::opts_chunk$set(echo = TRUE, eval = F, message = FALSE, warning = FALSE, fig.keep='none')
library(printr)
```

# Introduction

This tutorial introduces the basic model fitting process as implemented in [`brms`](https://paul-buerkner.github.io/brms/). This package provides an interface to fit Bayesian generalized (non-)linear multivariate multilevel models using Stan. Lucky for us, the formula syntax is very similar to that of the package `lme4` to provide a familiar and simple interface for performing regression analyses. So if you are familiar with fitting multilevel models in `lme4`, using `brms` will be a no-brainer. The interesting part is when we work with the posterior and add prior information to the models. 

# Preparation

## Loading packages

For this tutorial, we will load a number of different packages. First, we are going to load the `tidyverse` package set to have convenient data wrangling and visualization functions available. **Attention:** Always load the `tidyverse` before loading `brms`. For some weird reason, vice versa will break the functionality of the `tidyverse`. Second, we thus load the package `brms`, which is of course the core package for fitting Bayesian regression analysis. We are also going to load the packages `insight` and `bayestestR`, which both stem from the [`easystats`](https://easystats.github.io/easystats/) package set. The former provides functions to extract information from brms objects. The latter provides functions for Bayesian inferences. We also are going to load the package `bayesplot()`, which provides good functions for MCMC diagnostic checks. Finally, we are going to load the handy package `psych`, simply because it helps us understanding our data better.  

```{r}
library(tidyverse)
library(brms)
library(insight)
library(bayestestR)
library(bayesplot)
library(psych)
```


## Loading and recoding the data

We are going to use data by Masur, DiFranzo & Bazarova (2021), which we encountered in the examples during the lecture. The data was collected during an online experiment in 2019. The experiment was a 3x3 between-subject design that manipulated the prevaling *collective norm* in relation to visual disclosure. Participants were randomized into 9 different groups and confronted with a realistic, but controlled social media feed of ~50 posts. The feeds different with regard to a) the number of posts that featured a visual disclosure, i.e., whether the (fake) user showed his/her face in the post (0%, 20%, 80%), and b) the number of profile pictures that featured a visual disclosure, i.e., whether the (fake) user showed his/her face in the profile picture (0%, 20%, 80%). After going and interacting with the feed, participants completed a post-survey that included measures to assess norm perceptions, intention to disclose oneself visually as well as some control and potential moderators. 

We published the data set openly on the Open Science Framework (osf.io). The function `read.csv` can actually load such data sets directly into R (without technically downloading it). We directly transform the data set into a `tibble`, which is easier to handle compared to the standard `data.frame`. 

```{r}
d <- read.csv("https://osf.io/kfm6x/download", header = TRUE, na.strings = "NA") %>% 
  as_tibble
head(d)
```

In this tutorial, we are only going to focus on a subset of the variables. More specifically, we are going to investigate the effect of the (1) manipulations on the (2) norm perceptions and the (3) intention to disclose oneself. We will further add (4) the disposition for critical thinking and some (4) socio-demographics as control variables. In the code below, I am simply selecting the relevant variables based on their names (the variables conveniently have been named in a straight-forward manner) and pass those to the function `rowMeans`, which automatically computes the mean index. As a final step, we select only the variables that we are going to use in our analysis. 

```{r}
# Check variable naming
names(d)

# Recoding and selecting relevant variables
d <- d %>% 
  mutate(id = 1:n(),
         perceptions = rowMeans(d %>% select(contains("NOR"), -norm)),
         behavior = rowMeans(d %>% select(contains("BEH"))),
         crit_thinking = rowMeans(d %>% select(contains("CRI_THI"))), na.rm = T) %>%
  select(id, age, gender, edu, condition, norm, profile, perceptions, behavior, crit_thinking)

# New data set
head(d)
```


## Understanding the data set

Let's quickly run some descriptive analysis to better understand the data set. 

```{r}
# Data set dimensions
dim(d)

# Distribution across experimental groups
d %>%
  group_by(norm, profile) %>%
  tally %>%
  mutate(prop = n/nrow(d))

# Socio-demographics
describe(d$age)
table(d$gender)
table(d$edu)

# Small recoding
d <- d %>%
  mutate(gender = ifelse(gender == 2, NA, gender),
         edu = edu-1)
```

As we can see, we have 677 participants and 10 variables in the data set. Participants are fairly equally distributed across all nine conditions, which include further at least 71 participants. 


# Simple Regression Analysis

## Fitting a simple regression with `brms`

To run a (first) Bayesian simple regression analysis, we basically do the exact same thing as if we would use `lm()` to fit a Frequentist simple regression analysis. We only substitute `lm()` with `brm()`. The actually fitting process will take a bit longer that you might be used to as MCMC chains have to be simulated. 

Likewise, we can use the summary function to produce a first informative overview of the results. To keep it simple, we are going to investigate the relationship between two of the post-manipulation measures: Norm perceptions (perceptions) and intention to disclose oneself (behavior). The assumption is that the more people perceive the majority to disclose themselves, the more likely they are to disclose themselves. 

```{r}
m1 <- brm(behavior ~ perceptions, data = d)
summary(m1)
```

As we can see, the resulting summary even looks similar to what we would expect from a Frequentist analysis. Yet, it is important to understand that the table under *Population-Level Effects* are not point estimates, but summaries of posterior distributions. The "estimate" is in fact the median of this distribution and the "l-95% CI" and the "u-95% CI" are the 95% highest density intervals (HDI). With the *Rhat* values, we also get a first idea about whether the MCMC chains converged (remember: values close to 1 suggest convergence). 

Before we dive deeper into how we can adjust and change the defaults settings, let's quickly inspect the posterior distribution of our main effect visually. First, let's extract the posterior draws from the `brmsfit`-object. We can do so in various ways, but the `brms`-default is using the function `as_draws_df()`. 


```{r}
(posterior <- as_draws_df(m1))
```

As we can see, we get a data set with 4000 rows. The columns represent the different parameter that were estimated (Intercept, b, sigma). We further gain information about the chosen (default) prior and learn that 4 chains were created. So we can deduce that each chains contains 1000 draws. These are the `brms` defaults - we will learn later how to adjust these to fit our needs. For now it is sufficient to now that we basically have 4 posterior draws per parameter with 1000 values (combined 4000). 

We can now use this data set and visualize the posterior distribution for our effect of interest. The simplest way would be to simple plot a histogram. An alternative is to plot these draws as regression lines in a standard scatter plot. Here, we better sample from our posterior as 4000 lines would hardly fit on the scatterplot. 

```{r}
# Histogram
posterior %>%
  ggplot(aes(x = b_perceptions)) +
  geom_histogram(color = "white", fill = "steelblue") +
  theme_classic() +
  labs(x = "Effect of norm perceptions on intention to disclose")

# Regression lines
posterior %>%
  sample_n(size = 200) %>%
  ggplot() +
  geom_point(data = d,
             aes(x = perceptions, y = behavior),
             size = 1, alpha = .5) +
  geom_abline(aes(intercept = b_Intercept, 
                  slope = b_perceptions, 
                  group = .draw),
              color = "grey50", size = 1/4, alpha = .1) +
  theme_classic()
```

Although you can technically run any analysis like this, the real benefit of using Bayesian analysis comes with adjusting priors, investigating the convergence of the MCMC draws in more detail and using the posterior in different, more tailored-to-the-hypothesis/interest type of way. 

## Setting priors

In a first step, we may want to make the priors more informative as we probably have some subjective belief about our effect of interest that could be modeled as a distribution and included in the model. But what priors do we actually have to specify in this simple example? The `get_prior()` function can help here:

```{r}
get_prior(formula = behavior ~ perceptions, data = d)
```

When you run `get_prior()`, a table is returned with a few columns. Each row corresponds to  one model parameter. For now, we focus on the first three (prior, class, coef).

- *prior*: tells you which prior probability distributions are set as default by brms. For our model, the first two default priors are (flat), i.e. uniform distributions (all values are equally probable). The other two priors are Student-t distributions. This is what we will want to change to something more appropriate.

- *class:* indicates the type of model parameter: Intercept is the model intercept, b indicates a predictor term, and sigma the model’s residual standard deviation.

- *coef:* indicates to which model predictor the prior is assigned. In our model above, there’s only one predictor (perceptions), and that’s what you can see in the output.

Now, we want to change these priors to more appropriate prior assumptions. For our model, we need three priors: for the Intercept, the effect (b) and for sigma. We can simply store them as vector (`c()`) and define them with the function `prior()`.  Within the function, we need to specify the prior in "Stan" language and also specify to which class the prior belongs to. We probably have a more intuitive feeling for standardized effect sizes, so we specifiy the priors in terms of Cohen's *r*. Yet, we need to bear in mind that we then also have to standardize our coefficients before we refit the model.  

```{r}
# Quick visual check for my prior assumption about the effect
x <- seq(-1, 1, by = .001)
y <- dnorm(x, .2, 1)
ggplot(NULL, aes(x, y)) +
  geom_line()

# Quick visual check for the standard deviation/varince (= sigma)
library(extraDistr)
x <- seq(0, 10, by = 0.1)
y <- dhcauchy(x, sigma = 1)
ggplot(NULL, aes(x, y)) +
  geom_line()

# Specifiying priors
my_priors <- c(
  prior(normal(0, 10), class = "Intercept"),
  prior(normal(0.2, 1), class = "b"),
  prior(cauchy(0, 1), class = "sigma")
)
```

Now, we can use this vector and add it to the model.

```{r}
# Standardization to align with priors
d <- d %>%
  mutate(perceptions_std = scale(perceptions),
         behavior_std = scale(behavior))

# Model fitting
m2 <- brm(formula = behavior_std ~ perceptions_std, 
          prior = my_priors,
          chains = 2,        # only to reduce computation time, should be 4 at least
          iter = 12000,      # increase number of draws for more precise estimation
          warmup = 2000,     # set warmup
          sample_prior = T,  # This ensure that we can also draw from the prior (helps with visualization)
          data = d)

# Results
summary(m2)
```

The median effect size of the relationship between perceptions and behaviors is b = .55 (95% HDI[.49, .62]). Did anyone use different priors? What are your results?

Let's quickly check how the data updated our prior assumption. 

```{r}
# Get prior and posterior samples
posteriors <- get_parameters(m2) %>% as_tibble %>%
    mutate(type = "posterior")
priors <- prior_draws(m2) %>% as_tibble %>% 
    magrittr::set_colnames(names(posteriors)) %>%
    mutate(type = "prior")

# Plot
bind_rows(posteriors, priors)%>%
  ggplot(aes(x = b_perceptions_std, fill = type)) +
  geom_density(alpha = .8) +
  theme_classic() +
  xlim(-1, 1) +
  labs(x = expr(theta), y = "", fill = "") +
  scale_fill_manual(values = c("steelblue", "grey"))
```


# Convergence diagnostics

Before we turn to various approaches to Bayesian inference, we need to check more carefully whether our MCMC chains actually converged. We do so visually as well as by inspecting some statistics. 

## Trace plots

First, we investigate so-called trace plots, which show the random walk of the algorithm and check whether the chains overlap sufficiently. It should look like a somewhat "smooth caterpillar". 

```{r}
draws <- as_draws_df(m2)
mcmc_trace(draws, 
           pars = c("b_Intercept", "b_perceptions_std", "sigma"), 
           n_warmup = 0,
           facet_args = list(nrow = 2))
```

Based on these plots, we are not concerned that the chains did not converge. This is how these trace plots should look like. 


## Rhat

Next, we extract the Rhat values, a parameter for the variance between the chains in relation to the variance within the chain. It should be close to 1 to suggest convergence. 

```{r}
rhat(m2) %>%
  as_tibble(rownames = "par")
```

Again, no problem at all. 

## Autocorrelation

Finally, we check whether there is strong auto-correlation in the iterations of the chains. It looks like there is a bit of auto-correlation in the first chain. We can refit the model using thinning (that is using only every 4th draw) and check whether this improves the result. 

```{r}
mcmc_acf(draws, 
         pars = c("b_Intercept", "b_perceptions_std", "sigma"),
         lags = 20)

# Refit with thinning
m2b <- brm(formula = behavior_std ~ perceptions_std, 
          prior = my_priors,
          chains = 2,       
          iter = 12000,      
          warmup = 2000,     
          thin = 4,
          data = d)

drawsb <- as_draws_df(m2b)
mcmc_acf(drawsb, 
         pars = c("b_Intercept", "b_perceptions_std", "sigma"),
         lags = 20)
```

Now the auto-correlation plot look more like they should. 


# Bayesian Inference

## Highest Density Interval (HDI)

The highest density intervals (sometimes also HCR = Highest density region or HCI = Highest credibility interval) indicates which points of a distribution are most credible/probable. We can simply compute different HDIs with the function `hdi()` from the `bayestestR` package. We can also simply plot the output. 

```{r}
# Estimating hdis
(hdis <- hdi(draws$b_perceptions_std, ci = c(.90, .95, .99)))

# Plotting hdis
plot(hdis)
```

Based on the HDIs we can do simple Null-Hypothesis-Testing, by checking wether the NULL (whether exactly zero or any other value) is included in the HDI.


## Region of Practical Equivalence (ROPE)

Testing whether something is different than zero is often not very meaningful. We can hence also test whether the HDI lies outside a certain region that we assume to be equal to our NULL (e.g., a boundary around zero). The interesting aspect of this, that we can use a ROPE in different ways:

a) testing whether the effect is likely to be outside of a certain range
b) testing whether the effect is likely to be inside of a certain range (aking to equivalence testing)

```{r}
# Computing percentage within rope for a potential null = 0 region
rope(m2, range = c(-.1,.1), ci = c(.9, .95))

# Computing percentage within rope for a potential null = large effect region
rope(m2, range = c(.4,.6), ci = c(.9, .95))
```


Of course, we can also plot a ROPE testing approach. 

```{r}
a <- plot(rope(m2, range = c(-.1, .1), ci = .95)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  xlim(-.2, .8) +
  labs(caption = "Note: ROPE [-.10, .10]") +
  theme_minimal() +
  theme(legend.position = "none")

b <- plot(rope(m2, range = c(.4, .6), ci = .95)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  xlim(-.2, .8) +
  labs(caption = "Note: ROPE [.40, .60]", title = "") +
  theme_minimal()

cowplot::plot_grid(a, b, axis = "l", align = "v", rel_widths = c(1.5,2))
```


## Probability of direction

As mentioned in the lecture, we can also simply compute the probability of the effect being in a certain direction. 

```{r}
dir <- p_direction(draws$b_perceptions_std, null = 0)
plot(dir)

# Just to exemplify a case that is less certain
plot(p_direction(rnorm(1000, 0.1, .2)))
```


## Posterior Predictive Check

As a final step, we can simulate replicated data under the fitted model and then compare these to the observed data. These so-called posterior predictive checks help to identify systematic discrepancies between real and simulated data.

```{r}
pp_check(m2, ndraws = 100) + theme_minimal()
```

As we can see here, the fitted model does not yet replicate the data well. The reason is somewhat straight-forward: We are looking at post-experimental measures who have been affected by the manipulations. So let's dive deeper into the data set! 


# Exercise

**Let's practice what we just learned!**

In the example above, we have investigated the relationship between two of the post-experiment measures. Of course, it is way more interesting to test the actual effect of the manipulations. Let us create a full regression model in which we include both manipulations as independent variables (perhaps even including their interaction) as well as all control variables that we may find useful to include. Similar to Frequentist linear regression, we can of course use a stepwise approach and start with simpler models first and then gradually add predictors. The function `bayes_R2()` allows to compute a Bayesian equivalent of $R^2$ and could thus be used to compare explanatory power. So in sum, engage with the following steps:

1. Build meaningful models to predict the intention to disclose
2. Specify meaningful priors if possible (if you leave out those where you have now assumptions, the model will use a weakly informative default). 
3. Extract posterior for the relevant effects:
    - Effect of the norm manipulation on intention to disclose
    - Effect of the profile manipulation on the intention to disclose
    - Interaction between both?
4. Test whether these effects lie outside of a meaningful ROPE (e.g., -.1 to .1 = larger than small effects)

```{r}
# Code here
```



